---
title: "The United States of Oreo"
output: 
---

# The United States of Oreo

In the year 2026 Nabisco launched a sweepstakes, challenging the world to guess how many Oreo cookies it will take to completely cover the US Landmass. This R Markdown Document exists to document methods of calculating the answer.

```{r}
# Loading Libraries 
library(tigris)
library(sf)
library(dplyr)
library(ggplot2)
library(spatstat.geom)
library(terra)

```

# Defining the U.S

For the boundaries of the US we use the US Census Bureaus tiger dataset at 500k resultion.

US Territories have been excluded. This calcuation will only include the 50 US States.

```{r}

us_states <- states()

exclude_stusps <- c('AS', 
                    'GU', 
                    'MP', 
                    'VI', 
                    'PR') # American Samoa, Guam, Mariana Islands, Virgin Islands, Puerto Rico

us_states <- us_states %>%
  filter(!STUSPS %in% exclude_stusps)

ggplot(data = us_states) +
  geom_sf() +
  theme_classic()

# Dissolve state boundaries

us_states <- us_states %>%
  summarize(geometry = st_union(geometry))

ggplot(data = us_states) +
  geom_sf() +
  theme_classic()

```

# Transform into an equal area projection for higher accuracy

```{r}
# Print current CRS
st_crs(us_states)

us_states <- st_transform(us_states, 5070)

ggplot(data = us_states) +
  geom_sf() +
  theme_classic()
```

# Calculate Needed Oreos

```{r}

#  Get Area of US
A_us = st_area(us_states)
print(paste0('Area of US: ', A_us, ' m^2'))

# Calculate Area of Oreo 

## 1.8 inches in Diameter: https://onesweetmama.com/how-big-is-an-oreo/
## A = pi * r2

d_oreo_in = 1.75 
r_oreo_in = d_oreo_in/2
A_oreo_in2 = pi * r_oreo_in^2

# Convert to m^2
in_to_m = 0.0254 
sq_in_to_sq_m = in_to_m ^2 

A_oreo_m2 = A_oreo_in2 * sq_in_to_sq_m
print(paste0('Area of an Oreo', A_oreo_m2, ' m^2'))

# Calculate Number of Oreoes needed 
needed_oreos = A_us / A_oreo_m2
needed_oreos
```

## Adjust for Open water

```{r}

nlcd_lc_path <- "/home/brendan/Documents/united_states_of_oreo/data/Annual_NLCD_LndCov_2023_CU_C1V0/Annual_NLCD_LndCov_2023_CU_C1V0.tif"
nlcd_lc <- rast(nlcd_lc_path)

# clip and project

# create frequency table
freq_table <- freq(nlcd_lc)

water_pixels <- freq_table[freq_table$value == 11, "count"]

pixel_area <- res(nlcd_lc)[1] * res(nlcd_lc)[2]

water_area_m2 <- water_pixels * pixel_area

water_area_m2


```

## Adjust for Roads

```{r}

nlcd_id_path <- "/home/brendan/Documents/united_states_of_oreo/data/Annual_NLCD_ImpDsc_2023_CU_C1V0/Annual_NLCD_ImpDsc_2023_CU_C1V0.tif"
nlcd_id <- rast(nlcd_id_path)

# clip and project

# create frequency table
freq_table <- freq(nlcd_id)

road_pixels <- freq_table[freq_table$value == 1, "count"]

pixel_area <- res(nlcd_id)[1] * res(nlcd_id)[2]

roads_area_m2 <- road_pixels * pixel_area

roads_area_m2


```

## Adjust for Trees

## Adjust for Buildings

construction-physics.com estimates the building Square footage in the US to be 340 billion square feet

<https://www.construction-physics.com/p/every-building-in-america-an-analysis>

```{r}

building_area_sq_ft <- 3.4e11
building_area_m2 <- building_area_sq_ft * 0.09290304

```

## Final answer

```{r}
A_us <- as.numeric(A_us)
adjusted_area = A_us - water_area_m2 - roads_area_m2 - building_area_m2

adjusted_oreo_count = adjusted_area / oreo_hex_area_m

adjusted_oreo_count


```

```{r}



```
